---
title: "Final 2024 Take Home for Time Series"
author: Jamiu Olamilekan Badmus
date: "12-06-2024"
output: pdf_document
subtitle: Prof Yasin Kursat Onder
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4)
knitr::opts_chunk$set(warning = FALSE)
```

## Data Preparation
### Load the necessary libraries for the analysis
```{r, include=TRUE}
library(readr)
library(dynlm)
library(vars)
library(car)
library(tseries)
library(lpirfs)
library(urca)
```
### Load the data
```{r}
data <- read_csv("C:/Users/Jamiu/OneDrive - univ-lille.fr/Desktop/EGEI Second Semester - Ghent University/Applied Econometrics II/Time series/Final Exam Take home 2024/exam_2024.csv")
spec(data)
```

### Data Preprocessing
```{r}
realoilprices <- ts(data$rpoil, start = c(1975, 2), frequency = 4)
gdpdeflator <- ts(data$p, start = c(1975, 2), frequency = 4)
realgdp <- ts(data$y, start = c(1975, 2), frequency = 4)
SHOCKS <- ts(data$'oil consumption demand shocks', start = c(1975, 2), frequency = 4)
```

## 1. Plot the data, test for stationary and transformation if necessary
### Plot the data
```{r}
plot(realoilprices, main = "Figure 1: Real Oil Prices", ylab = "Real Oil 
     Prices", xlab = "Year")
plot(gdpdeflator, main = "Figure 2: GDP Deflator", ylab = "GDP Deflator", 
     xlab = "Year")
plot(realgdp, main = "Figure 3: Real GDP", ylab = "Real GDP", xlab = "Year")
```

### Test for stationarity
```{r}
adfraw1 <- ur.df(realoilprices, type = c("trend"), lags = 4, selectlags = "AIC")
summary(adfraw1)
adfraw2 <- ur.df(gdpdeflator, type = c("trend"), lags = 4, selectlags = "AIC")
summary(adfraw2)
adfraw3 <- ur.df(realgdp, type = c("trend"), lags = 4, selectlags = "AIC")
summary(adfraw3)
```

### Transformation of the series
```{r}
lrealoilprices <- log(realoilprices)
inflation <- (diff(gdpdeflator)/gdpdeflator[-length(gdpdeflator)])*100
lrealgdp <- log(realgdp)
```

### Plot the transformed data
```{r}
plot(lrealoilprices, main = "Figure 4: Log of Real Oil Prices", ylab = "Log of Real Oil Prices", 
     xlab = "Year")
plot(inflation, main = "Figure 5: Inflation", ylab = "Inflation", xlab = "Year")
plot(lrealgdp, main = "Figure 6: Log of Real GDP", ylab = "Log of Real GDP",
     xlab = "Year")
```

### Test for stationarity of the transformed data 
### Level Stationarity testing
```{r}
adflv1 <- ur.df(lrealoilprices, type = c("trend"), lags = 4, 
                selectlags = "AIC")
summary(adflv1)
adflv2 <- ur.df(inflation, type = c("trend"), lags = 4, selectlags = "AIC")
summary(adflv2)
adflv3 <- ur.df(lrealgdp, type = c("trend"), lags = 4, selectlags = "AIC")
summary(adflv3)
```

### Compute First Differenced Series
```{r}
dlrealoilprices <- diff(lrealoilprices)
dinflation <- diff(inflation)
dlrealgdp <- diff(lrealgdp)
```

### First difference Stationarity testing
```{r}
adfdiff1 <- ur.df(dlrealoilprices, type = c("trend"), lags = 4, 
                  selectlags = "AIC")
summary(adfdiff1)
adfdiff2 <- ur.df(dinflation, type = c("trend"), lags = 4, selectlags = "AIC")
summary(adfdiff2)
adfdiff3 <- ur.df(dlrealgdp, type = c("trend"), lags = 4, selectlags = "AIC")
summary(adfdiff3)
```

### Plot the first differenced series
```{r}
plot(dlrealoilprices, main = "Figure 7: First Difference of Log of Real Oil Prices", 
     ylab = "First Difference of Log of Real Oil Prices", xlab = "Year")
plot(dinflation, main = "Figure 8: First Difference of Inflation", 
     ylab = "First Difference of Inflation", xlab = "Year")
plot(dlrealgdp, main = "Figure 9: First Difference of Log of Real GDP", 
     ylab = "First Difference of Log of Real GDP", xlab = "Year")
```

## 2. Estimate a VAR model with the data
### Combine the variables into a matrix 
```{r}
x = cbind(dlrealoilprices, dinflation, dlrealgdp)
colnames(x) <- c("dlrealoilprices", "dinflation", "dlrealgdp")
x <- x[-1, ]
```

### Determine the VAR lag length using AIC criterion
```{r}
VAR_lag <- VARselect(x, lag.max = 8, type = "const")
print(VAR_lag)
```

### Determine the VAR lag length, specifically for AIC
```{r}
optimal_lag_AIC <- VAR_lag$selection["AIC(n)"]
print(optimal_lag_AIC)
```

## 3. Estimate the VAR model with the optimal lag length
```{r}
VAR_model <- VAR(x, p = optimal_lag_AIC, type = "const")
summary_VAR_model <- summary(VAR_model)
print (summary_VAR_model)
```

### Impulse Response Function for horizon 8
```{r}
irf1 <- irf(VAR_model, impulse = "dlrealoilprices", response = "dlrealoilprices", 
            n.ahead = 8)
plot(irf1)
irf2 <- irf(VAR_model, impulse = "dlrealoilprices", response = "dinflation",
            n.ahead = 8)
plot(irf2)
irf3 <- irf(VAR_model, impulse = "dlrealoilprices", response = "dlrealgdp",
            n.ahead = 8)
plot(irf3)
```

## 4. Estimate Local projections for the four variables 
### Combine the variables into a matrix 
```{r}
x1 = cbind(dlrealoilprices, dinflation, dlrealgdp, SHOCKS)
colnames(x1) <- c("dlrealoilprices", "dinflation", "dlrealgdp", "SHOCKS")
x1 <- x1[-c(1, 2), ]
x1_ts <- ts(x1, start = c(1975, 4), frequency = 4)
x1_df <- as.data.frame(x1_ts)
```

### Estimate the LP model with the optimal lag length
```{r}
lp <- lp_lin(x1_df, lags_endog_lin = optimal_lag_AIC,  trend = 0, 
             shock_type = 1,confint= 1.96, hor=8, use_nw = TRUE)
plot(lp)
```

### LP IRF plot 
```{r}
linear_plots <- plot_lin(lp)
linear_plots[[4]] # IRF for dlrealoilprices
linear_plots[[8]] # IRF for dinflation
linear_plots[[12]] # IRF for dlrealgdp
linear_plots[[16]] # IRF for SHOCKS
```
